---
title: "Homework_2_NTTNguyen"
author: "Ngoc Thuy Tram (Tram) Nguyen"
date: "2024-09-06"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1
```{r}
library(faraway)
head(prostate)
```

Below is the output after fitting the full model (i.e., **lpsa** as the response and all other variables (lcavol, **lweight**, **svi**, **lbph**, **age**, **lcp**, **pgg45**, and **gleason**) as predictors).
```{r}
model_full <- lm(lpsa ~ ., data=prostate)
model_full_output <- summary(model_full)
model_full_output
```

### 1a
```{r}
model_full_output$adj.r.squared*100
```
Percentage of variation in the response explained by these predictors is calculated as **100%*Adjusted R-squared**, which is 62.34%.

### 1b
```{r}
cor_fitted_response <- cor(model_full$fitted.values, prostate$lpsa)
cor_squared <- cor_fitted_response^2
cor_squared
```
This square of correlation between the fitted values and the response value is the **Multiple R-squared** from the model output.

### 1c
```{r}
i <- which.max(model_full$residuals)
i
```
The observation that has the largest residual is case **i=69**.

### 1d
```{r}
mean_resid <- mean(model_full$residuals)
mean_resid
median_resid <- median(model_full$residuals)
median_resid
```
Mean of the residuals is **mean_resid=-1.766244e-17**.
Median of the residuals is **median_resid=-0.01700021**.

### 1e
```{r}
cor_resid_fitted <- cor(model_full$resid, model_full$fitted.values)
cor_resid_fitted
```
Correlation of the residuals with the fitted values is **cor_resid_fitted=1.235656e-16**.

### 1f
When all other predictors held constant, the difference in predicted (log) prostate specific antigen for a patient with seminal vesicle invasion (svi=1) compared to a patient without seminal vesicle invasion (svi=0) is the value of coefficient of **svi**, which is **0.766157** from the model output.


## Question 2
From the **fat** dataset I will subset **new_fat**, which only includes the predictors involved in this Question 2.
```{r}
head(fat)
new_fat <- fat[,!(names(fat) %in% c("siri", "density", "adipos", "free"))] 
new_fat
```

The `get_rse_rsquare` function below is the one helps to compute the residual standard error and R square of all models, where predictors are added one at a time respectively.
```{r}
get_rse_rsquare <- function(x) {
     rse_rsquare_summary <- data.frame()
     y <- names(x[1])
     
     for (i in 2:ncol(x)) {
          x_temp <- x[,1:i]
          formula <- as.formula(paste(y," ~ ."))
          lmod <- lm(formula, data=x_temp)
          lmod_rse <- summary(lmod)$sigma
          lmod_adj_rsquared <- summary(lmod)$adj.r.squared
          obs <- c(y, paste(names(x_temp[2:i]), collapse=", "), lmod_rse, lmod_adj_rsquared)
          rse_rsquare_summary <- rbind(rse_rsquare_summary, obs)
     }
     
     colnames(rse_rsquare_summary) <- c("Response", "Predictor(s)", "RSE", "Adjusted_R_squared")
     print(rse_rsquare_summary)
}
```     

Below is the summary of RSE and R-squared of all 13 models
```{r}
get_rse_rsquare(new_fat)
```

### 2a
Plot of trend between RSE and R-squared
```{r}
library(ggplot2)
ggplot(get_rse_rsquare(new_fat), aes(x=as.numeric(RSE), y=as.numeric(Adjusted_R_squared))) +
     geom_point() +
     geom_smooth(method = "lm", se=FALSE) +
     scale_x_continuous(labels = function(x) sprintf("%.2f", x)) +  # Format x-axis to 2 decimal places
     scale_y_continuous(labels = function(y) sprintf("%.2f", y)) +
     labs(x = "RSE", y = "Adjusted R-squared", title = "Trends in RSE vs Adjusted R-squared of all models") +
     theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

### 2b
Fit the full model with **brozek** as the response
```{r}
model_full_2 <- lm(brozek ~ ., data = new_fat)
summary(model_full_2)
```

The coefficient of extended biceps circumference (i.e., **biceps** variable) is **0.15280**. This can be interpreted that the increase of 1 unit of extended biceps circumference will lead to a change of +0.15280 unit of body fat percentages (i.e., **brozek** variable) when all other variables are held constant.

### 2c
Fit the full model with **log (brozek)** as the response
```{r}

new_fat <- new_fat[new_fat$brozek != 0,]
new_fat$log_brozek <- log10(new_fat$brozek)
new_fat <- new_fat[, !names(new_fat) == "brozek"]
model_full_3 <- lm(log_brozek ~ ., data = new_fat)
summary(model_full_3)
```
--> Errors because brozek has 0 --> log(brozek) in undefined