---
title: "Homework 1"
author: "Ngoc Thuy Tram (Tram) Nguyen"
date: "2024-08-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Before going to answer question 1, I will install the **faraway** package
```{r}
install.packages("faraway")
library(faraway)
head(fat)
summary(fat)
str(fat)
```
### Numerical Summary
```{r}
head(fat)
summary(fat)
str(fat)
```

The **fat** dataset has a total of 18 numerical variables, in which either **brozek** or **siri** variable is the *response* variable, and the other remaining 16 variables are *predictor* variable in linear regression.
Among 16 *predictor* variables, I think **age, weight, height, adipos, abdom** are most likely to have a relationship with body fat percentage, so I will just analyze/give comments on these variables, along with **brozek** and **siri**.

### Graphical Summary

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

### Comments
