---
title: "Homework 1"
author: "Ngoc Thuy Tram (Tram) Nguyen"
date: "2024-08-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Firstly, I will load the **faraway** package.
```{r}
library(faraway)
```


Next, I will produce an overview about the **fat** dataset
```{r}
str(fat)
```

The **fat** dataset has a total of 18 numerical variables across 252 observations, in which either **brozek** or **siri** variable is the *response* variable, and the other remaining 16 variables are *predictor* variable in linear regression.
Among 16 *predictor* variables, I think **weight, height, adipos, abdom** are most likely to have a relationship with body fat percentage, so I will just analyze/give comments on these variables, along with either **brozek** or **siri**. Here, I will choose **brozek**. To be easier, I will save **brozek, weight, height, adipos, abdom** as a new data frame named **new_fat**
```{r}
new_fat <- subset(fat, select = c(brozek, weight, height, adipos, abdom))
```

### Numerical Summary
```{r}
summary(new_fat)
```

### Graphical Summary
Histogram of each variable with density curve
```{r}
par(mfrow=c(2, 3), mar=c(4,4,4,4))
hist(new_fat$weight, prob = TRUE)
lines(density(new_fat$weight), col = 4, lwd = 2)
hist(new_fat$height, prob = TRUE, ylim = c(0, max(density(new_fat$height)$y)))
lines(density(new_fat$height), col = 4, lwd = 2)
hist(new_fat$adipos, prob = TRUE, ylim = c(0, max(density(new_fat$adipos)$y)))
lines(density(new_fat$adipos), col = 4, lwd = 2)
hist(new_fat$abdom, prob = TRUE, ylim = c(0, max(density(new_fat$abdom)$y)))
lines(density(new_fat$abdom), col = 4, lwd = 2)
hist(new_fat$brozek, prob = TRUE, ylim = c(0, max(density(new_fat$brozek)$y)))
lines(density(new_fat$brozek), col = 4, lwd = 2)
```

Scatter-plot between each predictor (**weight, height, adipos, abdom**) and the response variable (**brozek**) 
```{r}
par(mfrow=c(2, 2))
plot(new_fat$weight, new_fat$brozek)
plot(new_fat$height, new_fat$brozek)
plot(new_fat$adipos, new_fat$brozek)
plot(new_fat$abdom, new_fat$brozek)
```

### Comments
* There are no missing values in **new_fat**. However, **brozek** has "0" value, which is invalid value for human body fat percentage and thus, double check is needed for observations with "0" value for **brozek** variable to ensure data accuracy. In case correct values for such observations can't be identified, it may be considered to remove such observations out of the dataset.
* All five variables have outliers that should be checked to see if they are reasonable.
* Without outliers, all five variables seem to have a normal distribution.
* Scatter plots suggest that each of the variables **weight, height, adipos** has a positive correlation with **brozek**, so they may be alternative methods to predict body fat percentage. However, further model selection/validation is needed to see if **weight, height, adipos** altogether, or just a subset of them, can be used to predict **brozek**.
* No relationship was found between **height** and **brozek**.

## Question 2

Firstly, I will load the **faraway** package.
```{r}
library(faraway)
```

Next, I will produce an overview about the **gala** dataset
```{r}
str(gala)
```

### Numerical Summary
In this Question 2, I will try calculating numerical summary using `dplyr` and `tidyr` packages for 3 variables **Species, Endemics, Area, Elevation**. To be easier, I will generate a new data frame named **new_gala** that includes these 4 variables only.
```{r}
library(faraway)
library(dplyr)
library(tidyr)

new_gala <- subset(gala, select = c(Species, Endemics, Area, Elevation))
new_gala %>% summarise(across(where(is.numeric), .fns = 
                     list(Min = min,
                          Median = median,
                          Q25 = ~quantile(., 0.25),
                          IQR = IQR,
                          Q75 = ~quantile(., 0.75),
                          Mean = mean,
                          SD = sd,
                          Max = max))) %>%
  pivot_longer(everything(), names_sep='_', names_to=c('Variable', '.value'))
```

### Graphical Summary
Histogram of each variable
```{r}
par(mfrow=c(2, 2), mar=c(4,4,4,4))
hist(new_gala$Species, breaks=30)
hist(new_gala$Endemics, breaks=30)
hist(new_gala$Area, breaks=30)
hist(new_gala$Elevation, breaks=30)
```

Box-plot of each variable
```{r}
par(mfrow=c(2, 2), mar=c(4,4,4,4))
boxplot(new_gala$Species, main="Species")
boxplot(new_gala$Endemics, main="Endemics")
boxplot(new_gala$Area, main="Area")
boxplot(new_gala$Elevation, main="Elevation")
```

Scatter-plot between each predictor (**Area, Endemics, Elevation**) and the response variable (**Species**) 
```{r}
par(mfrow=c(2, 2))
plot(new_gala$Area, new_gala$Species)
plot(new_gala$Endemics, new_gala$Species)
plot(new_gala$Elevation, new_gala$Species)
```

### Comments
* Histogram and box-plot charts show that distribution of 3 variables **Species, Endemics, Area, Elevation** are skewed, with some high outliers. This suggests that some transformation to another scale (e.g. log scale) may be considered.
* Scatter plot suggest that number of **Species** tend to increase as **Elevation** and **Endemics** get higher. Same pattern is observed between  **Species** and **Area**; however, this pattern is not as clear as the two pairs above.